Policy Gradient Algorithms

### 2019-03-20 pge-mattis.tex
Improvements in actor-critic section

Obviously, we do not have the true $Q_\pi(s,a)$, thus we need to approximate it by $\hat{Q}_\pi(s,a)$. 
In all of the following sections, when we say that we sample an episode, we mean to draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
Following this procedure, it is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the true local optimum of $J(\theta)$.

@@

However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0, \forall s \in \mathbb{S}$. 
Given this assumption, the function estimator $f_w$ is limited to approximating an advantage function
\begin{equation}
	f_w(s_t, a_t) \equiv \hat{A}_{\pi}(s_t, a_t) = \hat{Q}_{\pi}(s_t, a_t) - \hat{V}_{\pi}(s_t),
	\label{eqn:adv}
\end{equation} 
which requires bootstrapping for $\hat{V}_\pi(s)$. 
If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
This would render the process biased. 
Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} and Konda et al. \cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}. 
We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}_\pi(s,a)$ obtained from roll-outs. 
However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
This comes from the realization, that we use only a subspace of the actual action value function in $f_w^\pi$, which is only a state value function. 
One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $A_\pi(s,a) = f_w^\pi(s,a)$, $V_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
\begin{align}
  Q_\pi(s,a) = A_\pi(s,a) + V_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V_\pi(s')ds', \\
  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1}).
\end{align}
which gives involves only linear equations to solve \cite{4863}.

@@

From \eqref{eqn:nat-grad}, \eqref{eqn:F}, and \eqref{eqn:J-equals-F}, it follows that
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) = F^{-1}_\theta \nabla_\theta J(\theta) = F_\theta^{-1} F_\theta w = w.
\end{equation}
Thus, this approach does not require an actual estimate of the Fisher information matrix, but only an estimate of $w$, with the update step according to $\theta_{k+1} = \theta_k + \alpha_k w$.

Peters et al. \cite{4863} present this idea and suggest LSTD($\lambda$)-Q, a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic (eNAC).

>>>>

Obviously, we do not have the true $Q_\pi(s,a)$, thus we need to approximate it by $\hat{Q}_\pi(s,a)$. 
In all of the following sections, when we say that we sample an episode, we mean to draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
Following this procedure, it is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w^\pi(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the true local optimum of $J(\theta)$.

@@

However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w^\pi(s,a)da} = 0, \forall s \in \mathbb{S}$. 
Given this assumption, the function estimator $f_w^\pi$ is limited to approximating an advantage function
\begin{equation}
	f_w^\pi(s_t, a_t) \equiv \hat{A}_{\pi}(s_t, a_t) = \hat{Q}_{\pi}(s_t, a_t) - \hat{V}_{\pi}(s_t),
	\label{eqn:adv}
\end{equation} 
which requires bootstrapping for $\hat{V}_\pi$. 
If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
This would render the process biased. 
Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} and Konda et al. \cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}. 
We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}_\pi(s,a)$ obtained from roll-outs. 
However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
This comes from the realization, that we use only a subspace of the actual action value function in $\hat{V}^\pi$, which is only a state value function. 
One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
\begin{align}
  \hat{Q}_\pi(s,a) = \hat{A}_\pi(s,a) + \hat{V}_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)\hat{V}_\pi(s')ds', \\
  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1}).
\end{align}
which gives involves only linear equations to solve \cite{4863}.

@@

From \eqref{eqn:nat-grad}, and \eqref{eqn:J-equals-F}, it follows that
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) = F^{-1}_\theta \nabla_\theta J(\theta) = F_\theta^{-1} F_\theta w = w.
\end{equation}
Thus, this approach does not require an actual estimate of the Fisher information matrix, but only an estimate of $w$, with the update step according to $\theta_{k+1} = \theta_k + \alpha_k w$.

Peters et al. \cite{4863} present this idea and suggest LSTD-Q($\lambda$), a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic (eNAC).
